---
title: "Royal Power Sentiment Analysis"
subtitle: "Centralization & Decentralization in Denmark 1560–1660"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```

## Requirements

Install these packages once before knitting:

```r
install.packages(c("pdftools", "tidytext", "dplyr", "tidyr",
                   "ggplot2", "readr", "stringr", "knitr"))
```

```{r libraries}
library(here)
library(pdftools)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(knitr)
```

---

## Configuration

> **Set both paths here before knitting.**

```{r config}
PDF_PATH_1636     <- here("data","KB 1636-60.pdf")             # <-- full path to your PDF
LEXICON_PATH  <- here("data", "dk_royal_power_lexicon_clean.csv")  # <-- full path to the lexicon CSV
```

---

## 1. Extract Text from PDF

The entire PDF is read and merged into a single document for analysis.

```{r extract}
pdf_pages_1636 <- pdf_text(PDF_PATH_1636)

cat("Pages found:", length(pdf_pages_1636), "\n")
cat("Approx. word count:", 
    pdf_pages_1636 %>% paste(collapse = " ") %>% str_count("\\S+"), "\n")

# Combine all pages into one document
full_text <- tibble(
  unit_id = 1L,
  text    = paste(pdf_pages_1636, collapse = " ")
)
```

---

## 2. Load the Lexicon

```{r lexicon}
lexicon <- read_csv(LEXICON_PATH, show_col_types = FALSE)
cat("Lexicon entries:", nrow(lexicon), "\n")

lexicon %>%
  count(dimension, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  kable(caption = "Lexicon coverage by dimension and sentiment")
```

---

## 3. Tokenise & Match

```{r tokenise}
tokens <- full_text %>%
  unnest_tokens(word, text)

cat("Total tokens:", nrow(tokens), "\n")

matched <- tokens %>%
  inner_join(lexicon, by = "word")

cat("Matched tokens:", nrow(matched), "\n")
cat("Unique matched words:", n_distinct(matched$word), "\n")
```

---

## 4. Overall Sentiment Scores

```{r scores}
scores <- matched %>%
  summarise(
    total_score        = sum(score, na.rm = TRUE),
    mean_score         = round(mean(score, na.rm = TRUE), 3),
    n_matches          = n(),
    n_centralization   = sum(dimension %in% c("centralization",
                                               "authority")),
    n_decentralization = sum(dimension %in% c("decentralization",
                                               "resistance")),
    n_loyalty          = sum(dimension == "governance"),
    n_resistance       = sum(dimension == "resistance"),
    n_royal_grace      = sum(dimension == "governance"),
    n_royal_punishment = sum(dimension == "coercion"),
    cent_decent_ratio  = round(n_centralization / (n_decentralization + 1), 2)
  )

scores %>% kable(caption = "Sentiment scores across the full text")
```

```{r scores-interpretation}
direction <- if (scores$total_score > 0) {
  "**centralizing** — language of royal authority dominates."
} else if (scores$total_score < 0) {
  "**decentralizing** — language of resistance and noble autonomy dominates."
} else {
  "**neutral** — centralizing and decentralizing language are balanced."
}
```

> The overall sentiment of the text is `r direction`  
> Centralization/Decentralization ratio: **`r scores$cent_decent_ratio`** (>1 = more centralizing)

---

## 5. Dimension Breakdown

```{r dim-breakdown}
dim_counts <- matched %>%
  count(dimension, sentiment, sort = TRUE)

dim_counts %>%
  kable(caption = "Matched tokens by dimension and sentiment")
```

```{r plot-dimensions, fig.cap="Lexicon dimension frequency across full text"}
matched |>
  count(dimension) %>%
  ggplot(aes(x = reorder(dimension, n), y = n, fill = dimension)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title    = "Lexicon dimension frequency",
    subtitle = "How often each thematic category appears in the text",
    x = NULL, y = "Token count"
  ) +
  theme_minimal(base_size = 13)
```

---

## 6. Centralizing vs. Decentralizing Vocabulary

```{r plot-cent-decent, fig.cap="Centralizing vs. decentralizing term counts"}
tibble(
  type  = c("Centralizing terms", "Decentralizing / resistant terms"),
  count = c(scores$n_centralization, scores$n_decentralization)
) %>%
  ggplot(aes(x = type, y = count, fill = type)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_fill_manual(values = c("Centralizing terms"              = "#264653",
                               "Decentralizing / resistant terms" = "#e9c46a")) +
  labs(
    title    = "Centralizing vs. Decentralizing vocabulary",
    subtitle = "Centralizing: centralization + authority | Decentralizing: decentralization + resistance",
    x = NULL, y = "Token count"
  ) +
  theme_minimal(base_size = 13)
```

---

## 7. Loyalty vs. Resistance

```{r plot-loyalty-resistance, fig.cap="Loyalty vs. resistance to the king"}
tibble(
  type  = c("Loyalty to king", "Resistance to king"),
  count = c(scores$n_loyalty, scores$n_resistance)
) %>%
  ggplot(aes(x = type, y = count, fill = type)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_fill_manual(values = c("Loyalty to king"    = "#2a9d8f",
                               "Resistance to king" = "#e76f51")) +
  labs(
    title = "Loyalty vs. Resistance to the king",
    x = NULL, y = "Token count"
  ) +
  theme_minimal(base_size = 13)
```

---

## 8. Top 30 Matched Words

```{r plot-top-words, fig.height=8, fig.cap="Most frequent lexicon words found in the text"}
matched %>%
  count(word, dimension, sort = TRUE) %>%
  slice_head(n = 30) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word, fill = dimension)) +
  geom_col() +
  labs(
    title = "Top 30 matched lexicon words",
    x = "Frequency", y = NULL, fill = "Dimension"
  ) +
  theme_minimal(base_size = 12)
```

---

## 9. Full Matched Token Table

All words from the text that matched the lexicon, with their scores and dimensions.

```{r matched-table}
matched %>%
  count(word, dimension, sentiment, score, sort = TRUE) %>%
  rename(frequency = n) %>%
  kable(caption = "All matched lexicon words and their frequencies")
```

---

## 10. Save Results

```{r save}
# Save all plots as PNG files

ggsave("plot_dimensions.png",
       matched %>%
         count(dimension) %>%
         ggplot(aes(x = reorder(dimension, n), y = n, fill = dimension)) +
         geom_col(show.legend = FALSE) +
         coord_flip() +
         labs(title = "Lexicon dimension frequency",
              subtitle = "How often each thematic category appears in the text",
              x = NULL, y = "Token count") +
         theme_minimal(base_size = 13),
       width = 10, height = 6, dpi = 300)

ggsave("plot_cent_vs_decent.png",
       tibble(
         type  = c("Centralizing terms", "Decentralizing / resistant terms"),
         count = c(scores$n_centralization, scores$n_decentralization)
       ) %>%
         ggplot(aes(x = type, y = count, fill = type)) +
         geom_col(width = 0.5, show.legend = FALSE) +
         scale_fill_manual(values = c("Centralizing terms"              = "#264653",
                                      "Decentralizing / resistant terms" = "#e9c46a")) +
         labs(title    = "Centralizing vs. decentralizing vocabulary",
              subtitle = "Centralizing: centralization + authority | Decentralizing: decentralization + resistance",
              x = NULL, y = "Token count") +
         theme_minimal(base_size = 13),
       width = 10, height = 6, dpi = 300)

ggsave("plot_loyalty_vs_resistance.png",
       tibble(
         type  = c("Loyalty to king", "Resistance to king"),
         count = c(scores$n_loyalty, scores$n_resistance)
       ) %>%
         ggplot(aes(x = type, y = count, fill = type)) +
         geom_col(width = 0.5, show.legend = FALSE) +
         scale_fill_manual(values = c("Loyalty to king"    = "#2a9d8f",
                                      "Resistance to king" = "#e76f51")) +
         labs(title = "Loyalty vs. resistance to the king",
              x = NULL, y = "Token count") +
         theme_minimal(base_size = 13),
       width = 10, height = 6, dpi = 300)

ggsave("plot_top_words.png",
       matched %>%
         count(word, dimension, sort = TRUE) %>%
         slice_head(n = 30) %>%
         mutate(word = reorder(word, n)) %>%
         ggplot(aes(x = n, y = word, fill = dimension)) +
         geom_col() +
         labs(title = "Top 30 matched lexicon words",
              x = "Frequency", y = NULL, fill = "Dimension") +
         theme_minimal(base_size = 12),
       width = 10, height = 8, dpi = 300)

cat("Saved: plot_dimensions.png\n")
cat("Saved: plot_cent_vs_decent.png\n")
cat("Saved: plot_loyalty_vs_resistance.png\n")
cat("Saved: plot_top_words.png\n")
```

---

*Lexicon built from Kancelliets Breve 1611–35 | Theme: royal power, centralization & decentralization in Denmark 1560–1660*
